\section{Computational Cost Analysis}
\label{sec:cost}

We now turn to the central question of this paper: what does each turbulence closure cost to evaluate in a GPU-accelerated solver?
All 21 models are implemented in the same solver, use the same spatial discretization (second-order finite differences on a staggered MAC grid), the same time integration (SSP-RK3), and the same Poisson solver (cuFFT-based direct solve).
All timing is measured on a single NVIDIA H200 GPU using the solver's internal instrumentation (one \texttt{TIMED\_SCOPE} per component per RK3 substep).
This eliminates confounding factors that arise when comparing models across different codes, compilers, or hardware.

\subsection{Experimental setup}
\label{sec:cost:setup}

We measure per-step cost on four representative flow configurations spanning 2D and 3D, trained and unseen geometries, and IBM and wall-bounded cases:

\begin{table}[htbp]
\centering
\caption{Flow configurations for cost measurement. All use skew-symmetric convection, RK3 time integration, and FFT-based Poisson solvers.}
\label{tab:cost_cases}
\begin{tabular}{llrrll}
\toprule
Case & Geometry & Grid & Cells & Poisson & Training \\
\midrule
Cylinder Re$=100$ & Circular cylinder (IBM) & $384 \times 288$ & 0.1M & FFT2D & Unseen \\
Hills Re$_H=5{,}600$ & Periodic hills (IBM) & $1{,}536 \times 768$ & 1.2M & FFT2D & Trained \\
Duct Re$_b=3{,}500$ & Square duct & $96^3$ & 0.9M & MG & Trained \\
Sphere Re$=200$ & Sphere (IBM) & $384 \times 256^2$ & 25.2M & FFT 3D & Unseen \\
\bottomrule
\end{tabular}
\end{table}

To ensure that all models operate on developed flow with realistic velocity gradients, turbulence fields, and eddy viscosity, we employ a three-phase measurement protocol:
\begin{enumerate}
    \item \textbf{SST warm-up} (\texttt{warmup\_model = sst}, $t_\mathrm{warmup} = 2$~s): The SST $k$--$\omega$ transport model drives the flow from uniform initial conditions to a developed state with nonzero $k$, $\omega$, and $\nu_t$ fields. All timing data from this phase is discarded.
    \item \textbf{Model stabilization} (\texttt{warmup\_steps} = 50): The target turbulence model replaces SST, and the solver runs for 50 additional time steps to allow the model's eddy viscosity field to equilibrate. Timing data is again discarded.
    \item \textbf{Measurement} (100 time steps): Timing data is recorded. This is the only phase that contributes to the reported cost.
\end{enumerate}
All four configurations use a \emph{fixed time step} within the measurement phase to ensure that every model executes the same number of Poisson solves, convection evaluations, and diffusion updates at the same $\Delta t$.
This isolates the turbulence model cost from any indirect effects on the time step size (Section~\ref{sec:cost:adaptive} addresses the adaptive-$\Delta t$ scenario separately).

\subsection{Infrastructure cost is model-independent}
\label{sec:cost:infrastructure}

A key finding of this study is that the cost of every solver component \emph{other than} the turbulence model evaluation is identical across all 15 models, to within measurement noise ($\pm 1$--$2\%$).
Table~\ref{tab:infrastructure_cost} reports the per-substep cost of each solver component, averaged over all 15 models.

\begin{table}[htbp]
\centering
\caption{Infrastructure cost per RK3 substep (3 substeps per solver step), averaged across all 15 turbulence models. Standard deviation across models is $<2\%$ in all cases. The FFT Poisson solver is a direct solve with no iterative convergence---its cost depends only on grid size, not on the turbulence model's $\nu_t$ field.}
\label{tab:infrastructure_cost}
\begin{tabular}{lrrrr}
\toprule
Component & Hills (74K) & Cylinder (111K) & Duct (885K) & Sphere (3.15M) \\
\midrule
Poisson solve     & 0.275 ms & 0.280 ms & 0.105 ms & 0.377 ms \\
Convection        & 0.021 ms & 0.021 ms & 0.096 ms & 0.265 ms \\
Diffusion         & 0.020 ms & 0.021 ms & 0.071 ms & 0.177 ms \\
IBM forcing       & 0.020 ms & 0.021 ms & ---      & 0.083 ms \\
Vel.\ correction  & ---      & ---      & 0.079 ms & ---      \\
\bottomrule
\end{tabular}
\end{table}

The Poisson solver dominates the infrastructure cost on 2D grids (hills, cylinder), while convection and diffusion become significant on 3D grids (duct, sphere).
The IBM forcing cost is negligible ($< 0.1$~ms) on all grids---the pre-computed weight-array multiplication adds minimal overhead to the element-wise velocity update.

The model-independence of the Poisson cost deserves emphasis.
Because we use a direct FFT-based solver (not an iterative multigrid), the Poisson solve performs exactly the same operations regardless of the turbulence model's eddy viscosity field.
There are no additional iterations, no convergence-rate variations, and no model-dependent cost.
This means that the \emph{entire} cost difference between models is contained in the turbulence evaluation itself---the tables below report this cleanly separable overhead.

\subsection{Per-step turbulence model cost}
\label{sec:cost:perstep}

Table~\ref{tab:turb_cost_all} reports the turbulence model evaluation cost (ms per solver step) for all 15 models on all four cases.
The total step cost equals the baseline step cost plus the turbulence cost plus a small overhead for $\nu_\mathrm{eff}$ computation and residual evaluation.

\begin{table}[htbp]
\centering
\caption{Turbulence model evaluation cost (ms/step) on NVIDIA H200, fixed $\Delta t$, 100 measured steps after warm-up + stabilization. Models sorted by sphere cost. The baseline step cost (no turbulence) is 1.89~ms (hills), 2.05~ms (cylinder), 2.64~ms (duct), 6.03~ms (sphere).}
\label{tab:turb_cost_all}
\begin{tabular}{llrrrr}
\toprule
Model & Category & Hills & Cyl & Duct & Sphere \\
      &          & (74K) & (111K) & (885K) & (3.15M) \\
\midrule
Baseline (laminar) & --- & 0 & 0 & 0 & 0 \\
$k$-$\omega$       & Transport & 0.010 & 0.010 & 0.010 & 0.010 \\
SST $k$-$\omega$   & Transport & 0.025 & 0.026 & 0.037 & 0.074 \\
EARSM-WJ           & Transport+NLEV & 0.026 & 0.028 & 0.038 & 0.077 \\
EARSM-GS           & Transport+NLEV & 0.025 & 0.028 & 0.038 & 0.078 \\
EARSM-Pope         & Transport+NLEV & 0.025 & 0.028 & 0.038 & 0.077 \\
GEP                & Algebraic & 0.025 & 0.027 & 0.037 & 0.075 \\
Mixing length      & Algebraic & 0.054 & 0.058 & 0.064 & 0.102 \\
\midrule
TBRF (1 tree)      & NN (tree) & 0.069 & 0.087 & 0.503 & 1.588 \\
MLP ($32^2$)       & NN (dense) & 0.454 & 0.637 & 4.525 & 15.830 \\
MLP-med ($64^2$)   & NN (dense) & 1.434 & 2.099 & 16.223 & 57.367 \\
TBNN-s ($32^2$)    & NN (tensor) & 2.074 & 3.012 & 13.318 & 46.223 \\
PI-TBNN-s ($32^2$) & NN (tensor) & 2.076 & 3.008 & 13.286 & 46.270 \\
TBNN ($64^3$)      & NN (tensor) & 4.372 & 6.451 & 40.743 & 143.563 \\
PI-TBNN ($64^3$)   & NN (tensor) & 4.377 & 6.450 & 40.663 & 144.313 \\
\bottomrule
\end{tabular}
\end{table}

Several key observations emerge from Table~\ref{tab:turb_cost_all}:

\paragraph{Classical RANS models are negligible.}
All eight classical models (baseline through mixing length) cost less than 0.1~ms per step on every grid.
The relative overhead ranges from 0.2\% ($k$-$\omega$ on sphere) to 2.9\% (mixing length on cylinder).
This confirms the conventional wisdom that the turbulence model is not the computational bottleneck for classical RANS---the Poisson solve, convection, and diffusion dominate.

\paragraph{TBRF is the cheapest NN model by a wide margin.}
The TBRF with a single tree per basis coefficient costs 0.07--1.6~ms per step, corresponding to an overhead of 3--28\% relative to baseline.
This is an order of magnitude cheaper than any dense NN model.
The GPU tree traversal kernel, which evaluates all 10 basis coefficients in a single fused pass over the cell array, achieves this low cost through memory-bandwidth-bound execution with minimal arithmetic intensity.

\paragraph{TBNN and PI-TBNN have identical cost.}
The physics-informed TBNN (PI-TBNN) matches the standard TBNN within measurement noise at every grid size.
The realizability penalty that distinguishes PI-TBNN from TBNN is applied only during training (as an additional loss term); at inference time, both architectures perform the same forward pass through the same network.
This confirms that physics-informed training adds zero inference overhead.

\paragraph{NN cost scales linearly with grid size.}
Figure~\ref{fig:nn_scaling} (TODO) shows the scaling of turbulence model cost with cell count.
All dense NN models (MLP, TBNN) scale linearly---each cell's forward pass is independent (embarrassingly parallel matrix-vector multiply).
The TBRF scales sub-linearly, reflecting the memory-bandwidth-bound nature of tree traversal versus the compute-bound nature of dense matrix multiplication.
Classical models also scale linearly but with a much smaller per-cell cost.

\paragraph{Network size determines cost, not architecture type.}
Comparing models with similar parameter counts reveals that architecture matters less than size:
\begin{itemize}
    \item MLP ($32^2$, 1.2K params): 15.8~ms on sphere
    \item TBNN-small ($32^2$, 1.4K params): 46.2~ms on sphere
\end{itemize}
The TBNN-small is $2.9\times$ more expensive than the MLP despite similar parameter counts because the TBNN evaluates 10 tensor basis coefficients (10 outputs + 10 basis tensor computations) versus the MLP's single scalar output.
Within the same architecture family, cost scales roughly linearly with parameters:
\begin{itemize}
    \item MLP ($32^2$, 1.2K params) $\to$ MLP-med ($64^2$, 4.4K params): $3.6\times$ cost increase ($3.7\times$ parameter increase)
    \item TBNN-small ($32^2$, 1.4K params) $\to$ TBNN ($64^3$, 9.4K params): $3.1\times$ cost increase ($6.7\times$ parameter increase, sub-linear)
\end{itemize}

\subsection{Relative overhead}
\label{sec:cost:overhead}

Table~\ref{tab:overhead} expresses the turbulence model cost as a percentage of the total solver step, revealing how the overhead depends on grid size.

\begin{table}[htbp]
\centering
\caption{Turbulence model cost as percentage of total solver step time, NVIDIA H200. On 3D grids (duct, sphere), NN models dominate step cost; on 2D grids (hills, cylinder), the overhead is smaller because the base solver cost per cell is proportionally higher.}
\label{tab:overhead}
\begin{tabular}{lrrrr}
\toprule
Model & Hills & Cyl & Duct & Sphere \\
\midrule
$k$-$\omega$ & 0.5\% & 0.5\% & 0.4\% & 0.2\% \\
SST & 1.2\% & 1.3\% & 1.3\% & 1.2\% \\
EARSM (avg) & 1.3\% & 1.4\% & 1.4\% & 1.2\% \\
GEP & 1.3\% & 1.3\% & 1.4\% & 1.2\% \\
Mixing length & 2.7\% & 2.8\% & 2.4\% & 1.7\% \\
\midrule
TBRF-1t & 3.4\% & 4.2\% & 15.7\% & 20.6\% \\
MLP & 19.0\% & 23.7\% & 62.7\% & 72.4\% \\
MLP-med & 42.4\% & 50.3\% & 85.8\% & 90.3\% \\
TBNN-small & 51.0\% & 58.8\% & 83.1\% & 88.3\% \\
TBNN & 68.9\% & 75.3\% & 93.7\% & 95.9\% \\
PI-TBNN & 68.9\% & 75.3\% & 93.7\% & 95.9\% \\
\bottomrule
\end{tabular}
\end{table}

On the 3D sphere grid (3.15M cells), the TBNN consumes 96\% of the total step time---the solver infrastructure (Poisson, convection, diffusion, IBM) accounts for only 4\%.
Even the smallest TBNN variant (TBNN-small, $32^2$ hidden layers) dominates at 88\%.
The MLP with $32^2$ hidden layers reaches 72\%, and only the TBRF remains practical at 21\%.

The overhead is systematically lower on 2D grids because the solver's per-cell infrastructure cost is proportionally higher (the FFT2D Poisson solve is relatively more expensive per cell on 2D grids than FFT3D is on 3D grids, due to cuFFT's batching efficiency at larger problem sizes).

\subsection{Adaptive time step: wall time per physical time}
\label{sec:cost:adaptive}

The fixed-$\Delta t$ analysis above isolates the per-step model overhead.
In production, however, the adaptive time step creates a second cost mechanism: models that produce large $\nu_t$ reduce the diffusive CFL limit, forcing smaller $\Delta t$ and more time steps to simulate the same physical time.
Table~\ref{tab:wall_time} reports the total wall time to simulate a fixed interval of physical time (2~seconds) with adaptive $\Delta t$, using the same three-phase warm-up protocol.

\begin{table}[htbp]
\centering
\caption{Wall time (seconds) to simulate 2~s of physical time with adaptive $\Delta t$ on NVIDIA H200, after SST warm-up. The step count varies because each model's $\nu_t$ field determines the CFL-limited $\Delta t$. The ``slowdown'' column is relative to baseline.}
\label{tab:wall_time}
\begin{tabular}{lrrrrrr}
\toprule
& \multicolumn{3}{c}{Cylinder Re$=100$ (111K)} & \multicolumn{3}{c}{Duct Re$_b=3{,}500$ (885K)} \\
\cmidrule(lr){2-4} \cmidrule(lr){5-7}
Model & Wall (s) & Steps & Slow & Wall (s) & Steps & Slow \\
\midrule
Baseline      & 0.28  & 174   & $1.0\times$  & 0.79  & 298   & $1.0\times$ \\
SST           & 0.87  & 482   & $3.1\times$  & 0.84  & 298   & $1.1\times$ \\
EARSM-WJ     & 0.37  & 196   & $1.3\times$  & 0.85  & 298   & $1.1\times$ \\
TBRF-1t       & 8.21  & 4659  & $29\times$   & 0.96  & 298   & $1.2\times$ \\
MLP           & 11.01 & 4654  & $39\times$   & 125.5 & 17357 & $159\times$ \\
TBNN-small    & 22.87 & 4663  & $82\times$   & 4.84  & 298   & $6.1\times$ \\
\bottomrule
\end{tabular}
\end{table}

The adaptive-$\Delta t$ results reveal a qualitatively different cost landscape than the fixed-$\Delta t$ analysis:

\paragraph{MLP crushes the time step on all geometries.}
On the cylinder, the MLP takes 4{,}654 steps versus baseline's 174 ($27\times$ more steps) because its large $\nu_t$ prediction reduces $\Delta t$ from $1.2 \times 10^{-2}$ to $4.3 \times 10^{-4}$.
Combined with the 47\% per-step model overhead, this yields a total slowdown of $39\times$.
On the duct, the effect is even more dramatic: the MLP takes 17{,}357 steps ($58\times$ more) for a total slowdown of $159\times$.
This systematic over-prediction of eddy viscosity is a generalization failure that imposes a severe practical cost penalty even when the model nominally ``runs'' without crashing.

\paragraph{TBRF-1t is penalized by $\Delta t$ on the cylinder but not on the duct.}
On the cylinder (an unseen geometry), the TBRF-1t takes 4{,}659 steps ($27\times$ baseline), producing a $29\times$ total slowdown despite its low per-step cost.
On the duct (a trained geometry), the TBRF-1t takes the same 298 steps as baseline, yielding only $1.2\times$ total slowdown.
This demonstrates that the $\Delta t$ penalty is geometry-dependent and reflects the model's accuracy, not its computational architecture.

\paragraph{Two distinct cost mechanisms.}
The total cost of a turbulence closure in production is:
\begin{equation}
    C_\mathrm{total} = \underbrace{C_\mathrm{step}}_{\text{per-step overhead}} \times \underbrace{N_\mathrm{steps}}_{\text{adaptive } \Delta t \text{ penalty}}
\end{equation}
The fixed-$\Delta t$ tables isolate $C_\mathrm{step}$; the adaptive-$\Delta t$ table captures $C_\mathrm{total}$.
A model can be cheap per step but expensive in total if it produces excessive $\nu_t$ (MLP on cylinder), or expensive per step but cheap in total if its $\nu_t$ is well-calibrated (TBNN-small on duct).
The Pareto-optimal choice depends on whether the application is accuracy-limited or cost-limited.

\subsection{Component breakdown}
\label{sec:cost:breakdown}

To visualize where the computational time is spent, Table~\ref{tab:breakdown_sphere} provides a complete component breakdown for the sphere case (3.15M cells), which has the highest absolute costs and the clearest separation between components.

\begin{table}[htbp]
\centering
\caption{Complete per-step cost breakdown for sphere Re$=200$ (3.15M cells), NVIDIA H200, fixed $\Delta t = 0.015$. RK3 calls convection, diffusion, Poisson, IBM, and velocity correction 3$\times$ per step; turbulence is called once.}
\label{tab:breakdown_sphere}
\begin{tabular}{lrrrrrrr}
\toprule
Model & Total & Turb & Poisson & Conv & Diff & IBM & Other \\
      & (ms)  & (ms) & (ms)    & (ms) & (ms) & (ms)& (ms) \\
\midrule
Baseline        &  6.03 &  0     & 1.13 & 0.80 & 0.53 & 0.25 & 3.32 \\
SST             &  6.24 &  0.07  & 1.13 & 0.80 & 0.53 & 0.25 & 3.46 \\
EARSM-WJ       &  6.32 &  0.08  & 1.13 & 0.80 & 0.53 & 0.25 & 3.53 \\
TBRF-1t         &  7.72 &  1.59  & 1.13 & 0.80 & 0.53 & 0.25 & 3.42 \\
MLP ($32^2$)    & 21.98 & 15.83  & 1.13 & 0.80 & 0.53 & 0.25 & 3.44 \\
TBNN-s ($32^2$) & 52.37 & 46.22  & 1.13 & 0.80 & 0.53 & 0.25 & 3.44 \\
TBNN ($64^3$)   &149.76 &143.56  & 1.13 & 0.80 & 0.53 & 0.25 & 3.49 \\
\bottomrule
\end{tabular}
\end{table}

The Poisson solve (1.13~ms $= 3 \times 0.377$~ms/substep), convection (0.80~ms), and diffusion (0.53~ms) are perfectly constant across all models.
The IBM forcing (0.25~ms) is also constant.
The ``Other'' column captures the velocity correction, $\nu_\mathrm{eff}$ computation, residual evaluation, boundary conditions, and RK3 bookkeeping.
The \emph{only} component that varies is the turbulence model evaluation.

This clean decomposition is possible because the solver uses an FFT-based direct Poisson solve with no iterative convergence, and because the fixed-$\Delta t$ protocol ensures identical work in every other solver phase.
With an iterative Poisson solver (e.g., multigrid), the Poisson cost could vary between models if different $\nu_t$ fields produce differently conditioned pressure equations.

\subsection{Cost-accuracy tradeoff}
\label{sec:pareto}

The per-step cost data from Table~\ref{tab:turb_cost_all} provides the $x$-axis for the Pareto cost-accuracy plot.
The $y$-axis (a posteriori prediction accuracy) will be populated from the production a posteriori runs (Section~\ref{sec:aposteriori}).
The combination of per-step overhead (computational architecture) and adaptive-$\Delta t$ penalty (model accuracy) defines the full cost landscape that practitioners must navigate when selecting a closure.
