\documentclass[11pt]{article}

% --- Standard packages ---
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{natbib}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage{siunitx}
\usepackage{multirow}
\usepackage{bm}

% --- Custom commands ---
\newcommand{\todo}[1]{\textcolor{red}{\textbf{[TODO: #1]}}}
\newcommand{\Rey}{\mathrm{Re}}
\newcommand{\Reytau}{\Rey_\tau}
\newcommand{\bhat}{\hat{b}}
\newcommand{\Shat}{\hat{S}}
\newcommand{\Omhat}{\hat{\Omega}}
\newcommand{\nueff}{\nu_\mathrm{eff}}
\newcommand{\nusgs}{\nu_\mathrm{sgs}}
\newcommand{\nut}{\nu_t}

\title{Cost-Accuracy Tradeoffs for Neural Network Turbulence Closures\\in GPU-Accelerated Incompressible Flow Solvers}

\author{TBD}

\date{}

\begin{document}

\maketitle

\begin{abstract}
We systematically compare five data-driven turbulence closure architectures---multilayer perceptron (MLP), large MLP, tensor basis neural network (TBNN), physics-informed TBNN (PI-TBNN), and tensor basis random forest (TBRF)---against classical RANS models in a GPU-accelerated incompressible Navier--Stokes solver.
All models are trained on the same dataset \citep{McConkey2021} with the same case-holdout protocol and evaluated both \emph{a priori} (offline prediction accuracy) and \emph{a posteriori} (coupled CFD accuracy and wall-clock cost).
We provide a detailed computational anatomy of each closure, showing exactly which operations---matrix multiplies, activation functions, tree traversals, transport equation solves---dominate the cost.
On a \todo{GPU type} GPU, the TBNN achieves \todo{X}\% lower prediction error than $k$-$\omega$ SST at \todo{Y}\% additional computational cost per time step, while the smaller MLP adds only \todo{Z}\% overhead.
The TBRF yields the highest offline accuracy (validation RMSE of 0.064 vs.\ 0.085 for TBNN) but is \todo{N}$\times$ more expensive in the solver due to branch-heavy tree traversals and poor GPU cache utilization.
Physics-informed realizability penalties on the TBNN provide no benefit: the tensor basis architecture already constrains outputs to be near-realizable, and increasing the penalty weight degrades accuracy.
\todo{Add key a posteriori finding.}
These results provide practitioners with quantitative guidance for selecting turbulence closures that balance predictive accuracy against computational budget.
\end{abstract}

\input{sections/introduction}
\input{sections/methods_solver}
\input{sections/methods_closures}
\input{sections/methods_training}
\input{sections/results_apriori}
\input{sections/results_aposteriori}
\input{sections/results_cost}
\input{sections/discussion}
\input{sections/conclusions}

\appendix
\input{sections/appendix}

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
